---
title: "Trees_model_app_description"
author: "Jakub Tomaszewski"
date: "Saturday, July 26, 2014"
output: html_document
---

Introduction
----------------------------
This document contains technical information on the presented Shiny app.

*Trees_model_app* uses data from 'trees' dataset, which consists of 31 observations and 3 attributes describing size of the given tree: Girth, Height and Volume. 

User of this app can experiment with the linear models predicting volume of the given tree with one explanatory variable: Girth or Height.

Features
----------------------------

- Choice of the predictor used in the model.
- Control of the confidence level for confidence intervals on the plot.
- Number of observations in the data can be changed (i.e. to check how the width of confidence intervals depends on the size of data).
- Adjustment of the plot with checkboxes (show/hide options):
    1. Grid;
    2. Line fitted by the model;
    3. Curves corresponding to the borders of confidence intervals for the fitted values.

Code
----------------------------

Here is the code for handling user interface of this app:

```{r, results='hide'}
library(shiny)

shinyUI(
  navbarPage("Predictive model for the tree volume (based on 'trees' dataset)",
             tabPanel("Linear model and graph",
                      
                      sidebarLayout(
                        sidebarPanel(
                          sliderInput("n_obs",
                                      "Number of observations used for building the linear model:",
                                      min = 1,
                                      max = nrow(trees),
                                      value = nrow(trees),
                                      step = 1),
                          checkboxInput("grid","Show grid on the plot",TRUE),
                          selectInput("predictor1", "Predictor (visible on X-axis):",choices=c("Girth","Height")),
                          checkboxInput("trend","Show linear regression trend line based on the predictor chosen above"),
                          sliderInput("conf_level",
                                      "Choose the confidence level [%]:",
                                      min = 0,
                                      max = 99,
                                      value = 95,
                                      step = 1),
                          checkboxInput("conf","Show the confidence interval for the trend line")
                        ),
                        
                        mainPanel(
                          plotOutput("treePlot")
                        )
                      )
             ),
             tabPanel("Documentation",
                      mainPanel(
                        includeMarkdown("trees_model_app_description.md")
                      )
             )
            )
)
```

Code responsible for server-side calculations is listed below:

```{r, results='hide'}
library(shiny)
data(trees)
n <- nrow(trees)

Girth <- trees$Girth
Height <- trees$Height
Volume <- trees$Volume

shinyServer(function(input, output) {

  output$treePlot <- renderPlot({

   n_obs <- input$n_obs
   observations_indices <- 1:n_obs
    
    plot(x=trees[observations_indices,input$predictor1], y=trees[observations_indices,"Volume"],
         col='green',
         pch=19,
         cex=1.5,
         xlab=input$predictor1,
         ylab="Volume",
         main=paste0("Scatterplot of ",input$predictor1," and Volume"))
    if(input$grid==TRUE) grid()
    if(input$trend==TRUE){
      model <- lm(paste0("Volume ~ ",input$predictor1),data=trees)
      abline(model)
    }
    if(input$conf==TRUE){
      conf_x <- data.frame(seq(min(trees[,input$predictor1]), max(trees[,input$predictor1]), 0.1))
      colnames(conf_x) <- input$predictor1
      model <- lm(paste0("Volume ~ ",input$predictor1),data=trees)
      pred <- predict(model,
                      newdata=conf_x,
                      interval="confidence",
                      level=input$conf_level/100)
      lines(conf_x[,1], pred[,2], lty=2)
      lines(conf_x[,1], pred[,3], lty=2) 
    }
  })
})

```

I hope you enjoy this simple app.